Reflections on reflection for \\ intuitionistic set theories

### 2019-01-10 paper-reflection.tex
typo

\begin{lemma}\label{lemma:zf-smallstep}
Let~$\varphi(u_1,\ldots,u_n,x)$ be a formula in the language of set
theory with free variables as indicated. Then~$\ZF$ proves
\begin{multline*}
  {\qquad}\forall M\_
  \exists S \supseteq M\_
  \forall u_1,\ldots,u_n \in S\_ \\
  (\exists x\_ \varphi(u_1,\ldots,u_n,x)) \Longrightarrow
  (\exists x \in S\_ \varphi(u_1,\ldots,u_n,x)).{\qquad}
\end{multline*}
Furthermore:
\begin{enumerate}
\item We may suppose that~$S$ is transitive.
\item We may suppose that~$S$ is closed under subsets.
\item Given a finite list~$\varphi_1,\ldots,\varphi_s$ of formulas instead of
the single formula~$\varphi$, we may suppose that~$S$ bounds all of
the formulas~$\varphi_i$.
% XXX: better explanation
\end{enumerate}
\end{lemma}

@@

\begin{lemma}\label{lemma:izf-microstep}
Let~$\varphi(u_1,\ldots,u_n,x)$ be a formula in the language of set
theory with free variables as indicated. Then~$\IZF$ proves
\begin{multline*}
  {\qquad}\forall H\_
  \exists H' \supseteq H\_
  \forall u_1,\ldots,u_n \in H\_ \\
  \begin{array}{@{}rcl@{}}
  (\exists x\_ \varphi(u_1,\ldots,u_n,x)) &\Longrightarrow&
  (\exists x \in H'\_ \varphi(u_1,\ldots,u_n,x)) \quad\mathop{\wedge} \\
  (\forall x\_ \varphi(u_1,\ldots,u_n,x)) &\Longleftarrow&
  (\forall x \in H'\_ \varphi(u_1,\ldots,u_n,x)).\end{array}
  {\qquad}
\end{multline*}
Furthermore, we may assume that~$H'$ is transitive and closed under subsets;
and given a finite list~$\varphi_1,\ldots,\varphi_s$ of formulas instead of
the single formula~$\varphi$, we may suppose that~$H'$ has the displayed
property for each of the formulas~$\varphi_i$.
\end{lemma}

@@

\begin{lemma}\label{lemma:izf-smallstep}
Let~$\varphi(u_1,\ldots,u_n,x)$ be a formula in the language of set
theory with free variables as indicated. Then~$\IZF+\RRS$ proves
\begin{multline*}
  {\qquad}\forall M\_
  \exists S \supseteq M\_
  \forall u_1,\ldots,u_n \in S\_ \\
  \begin{array}{@{}rcl@{}}
  (\exists x\_ \varphi(u_1,\ldots,u_n,x)) &\Longrightarrow&
  (\exists x \in S\_ \varphi(u_1,\ldots,u_n,x)) \quad\mathop{\wedge} \\
  (\forall x\_ \varphi(u_1,\ldots,u_n,x)) &\Longleftarrow&
  (\forall x \in S\_ \varphi(u_1,\ldots,u_n,x)).\end{array}
  {\qquad}
\end{multline*}
Furthermore:
\begin{enumerate}
\item We may suppose that~$S$ is transitive.
\item We may suppose that~$S$ is closed under subsets.
\item Given a finite list~$\varphi_1,\ldots,\varphi_s$ of formulas instead of
the single formula~$\varphi$, we may suppose that~$S$ bounds all of
the formulas~$\varphi_i$.
% XXX: better explanation
\end{enumerate}
\end{lemma}

>>>>

\begin{lemma}\label{lemma:zf-smallstep}
Let~$\varphi(u_1,\ldots,u_n,x)$ be a formula in the language of set
theory with free variables as indicated. Then~$\ZF$ proves
\begin{multline*}
  {\qquad}\forall M\_
  \exists S \supseteq M\_
  \forall u_1,\ldots,u_n \in S\_ \\
  (\exists x\_ \varphi(u_1,\ldots,u_n,x)) \Longrightarrow
  (\exists x \in S\_ \varphi(u_1,\ldots,u_n,x)).{\qquad}
\end{multline*}
Furthermore: (1)~We may suppose that~$S$ is transitive. (2)~We may suppose that~$S$ is
closed under subsets. (3)~Given a finite list~$\varphi_1,\ldots,\varphi_s$
of formulas instead of the single formula~$\varphi$, we may suppose that~$S$
bounds all of the formulas~$\varphi_i$.
% XXX: better explanation
\end{lemma}

@@

\begin{lemma}\label{lemma:izf-microstep}
Let~$\varphi(u_1,\ldots,u_n,x)$ be a formula in the language of set
theory with free variables as indicated. Then~$\IZF$ proves
\begin{multline*}
  {\qquad}\forall H\_
  \exists H' \supseteq H\_
  \forall u_1,\ldots,u_n \in H\_ \\
  \begin{array}{@{}rcl@{}}
  (\exists x\_ \varphi(u_1,\ldots,u_n,x)) &\Longrightarrow&
  (\exists x \in H'\_ \varphi(u_1,\ldots,u_n,x)) \quad\mathop{\wedge} \\
  (\forall x\_ \varphi(u_1,\ldots,u_n,x)) &\Longleftarrow&
  (\forall x \in H'\_ \varphi(u_1,\ldots,u_n,x)).\end{array}
  {\qquad}
\end{multline*}
Furthermore: (1)~We may suppose that~$H'$ is transitive. (2)~We may suppose
that~$H'$ is closed under subsets. (3)~Given a finite
list~$\varphi_1,\ldots,\varphi_s$ of formulas instead of the single
formula~$\varphi$, we may suppose that~$H'$ has the displayed property for each
of the formulas~$\varphi_i$.
\end{lemma}

@@

\begin{lemma}\label{lemma:izf-smallstep}
Let~$\varphi(u_1,\ldots,u_n,x)$ be a formula in the language of set
theory with free variables as indicated. Then~$\IZF+\RRS$ proves
\begin{multline*}
  {\qquad}\forall M\_
  \exists S \supseteq M\_
  \forall u_1,\ldots,u_n \in S\_ \\
  \begin{array}{@{}rcl@{}}
  (\exists x\_ \varphi(u_1,\ldots,u_n,x)) &\Longrightarrow&
  (\exists x \in S\_ \varphi(u_1,\ldots,u_n,x)) \quad\mathop{\wedge} \\
  (\forall x\_ \varphi(u_1,\ldots,u_n,x)) &\Longleftarrow&
  (\forall x \in S\_ \varphi(u_1,\ldots,u_n,x)).\end{array}
  {\qquad}
\end{multline*}
Furthermore: (1)~We may suppose that~$S$ is transitive. (2)~We may suppose
that~$S$ is closed under subsets. (3)~Given a finite
list~$\varphi_1,\ldots,\varphi_s$ of formulas instead of the single
formula~$\varphi$, we may suppose that~$S$ bounds all of the
formulas~$\varphi_i$.
% XXX: better explanation
\end{lemma}
